---
title: "Covid 19 vaccinations"
output:
  flexdashboard::flex_dashboard:
    fig_keep: high
    cache_keep: high
---

```{r}
options(repos = "https://cran.rstudio.com/")


```


```{r}
install.packages("Rtools")
install.packages("readxl")
library("readxl")
```

```{r Working directory}

getwd()

```



```{r}
library(readxl)

pfizer <- "datasets//pfizer.xlsx"
pfizer <- read_excel(pfizer)

Moderna<- "datasets//moderna.xlsx"
Moderna<- read_excel(Moderna)

janssen <- "datasets//janssen.xlsx"
janssen <- read_excel(janssen)


```



```{r}

install.packages("tidyverse")
install.packages("dplyr")
library(tidyverse)
library(dplyr)
```



```{r}
install.packages("devtools")
library(devtools)
devtools::install_github('FIRST-Data-Lab/IDDA', force = TRUE)
```

# creating variables from Datasets
```{r}
State <- pfizer$Jurisdiction
Date <- pfizer$Week.of.Allocations
Pfizer.1st.Dose.Allocations <- pfizer$X1st.Dose.Allocations 
Pfizer.2nd.Dose.Allocations <- pfizer$X2nd.Dose.Allocations 
Moderna.1st.Dose.Allocations <- Moderna$X1st.Dose.Allocations
Moderna.2nd.Dose.Allocations <- Moderna$X2nd.Dose.Allocations 
Janssen.1st.Dose.Allocations <- janssen$X1st.Dose.Allocations 
All.Dose.Allocations <- 0.5 *(Pfizer.2nd.Dose.Allocations+Pfizer.1st.Dose.Allocations+Moderna.1st.Dose.Allocations + Moderna.2nd.Dose.Allocations)+Janssen.1st.Dose.Allocations
Cum.Allocation <-cumsum(All.Dose.Allocations)

```


#1.Data Exploration.

# find out the maximum length column, inorder to create a dataframe with equal number of records. As we know different vaccine datasets has different number of records.

```{r}
max_length <- max(c(length(Janssen.1st.Dose.Allocations),length(Moderna.1st.Dose.Allocations),length(Moderna.2nd.Dose.Allocations),length(Pfizer.1st.Dose.Allocations),length(Pfizer.2nd.Dose.Allocations)))
max_length
```


```{r}
lengths_vector <- c(length(Janssen.1st.Dose.Allocations), length(Moderna.1st.Dose.Allocations), length(Moderna.2nd.Dose.Allocations), length(Pfizer.1st.Dose.Allocations), length(Pfizer.2nd.Dose.Allocations))

if (length(lengths_vector) > 0) {
  max_length <- max(lengths_vector)
} else {
  # Handle the case where all vectors have length zero
  max_length <- 0
}


```


#creating a dummy vaccine_dataframe
```{r}
df_vaccine <- data.frame(
  Janssen.1st.Dose.Allocations = c(Janssen.1st.Dose.Allocations, rep(NA, max_length - length(Janssen.1st.Dose.Allocations))),
  Moderna.1st.Dose.Allocations = c(Moderna.1st.Dose.Allocations, rep(NA, max_length - length(Moderna.1st.Dose.Allocations))),
  Moderna.2nd.Dose.Allocations = c(Moderna.2nd.Dose.Allocations, rep(NA, max_length - length(Moderna.2nd.Dose.Allocations))),
  Pfizer.1st.Dose.Allocations = c(Pfizer.1st.Dose.Allocations, rep(NA, max_length - length(Pfizer.1st.Dose.Allocations))),
  Pfizer.2nd.Dose.Allocations = c(Pfizer.2nd.Dose.Allocations, rep(NA, max_length - length(Pfizer.2nd.Dose.Allocations))),
  State = c(State, rep(NA, max_length - length(State))),
  Date = c(Date, rep(NA, max_length - length(Date))),
  All.Dose.Allocations = c(All.Dose.Allocations, rep(NA, max_length - length(All.Dose.Allocations))),
  Cum.Allocation = c(Cum.Allocation, rep(NA, max_length - length(Cum.Allocation)))
)

```



```{r}

str(df_vaccine)

```



```{r}
population_data <- IDDA::pop.state

```


```{r}
population_data  <- population_data%>% rename_at('State', ~'name')

```



# removing the space from state column
```{r}
df_vaccine <- df_vaccine%>%
  mutate(name = sapply(State, gsub, pattern = " ", 
replacement = ""))

```




#creating a left_join using Dplyr
```{r}

df<- left_join(df_vaccine,population_data, by="name")

```




#creating Doseperpop Variable using Mutate

```{r}
library(dplyr)
df <- mutate(df, Doseperpop=Cum.Allocation/df$population)

```
#Making sure to drop all the NA values


```{r}


df <- df[!(df$State=="American Samoa" | df$State=="Palau" | df$State == "Guam" | df$State == "Mariana Islands" | df$State == "Marshall Islands" | df$State == "PuertoRico" | df$State == "Philadelphia" | df$State == "New York City" |df$State == "Puerto Rico" |df$State == "U.S. Virgin Islands" | df$State == "Chicago" |df$State == "Federal Entities" | df$State == "Micronesia"),] 

```


```{r}
str(df)
```



#2.Visualization

```{r}
install.packages("plotly")
library(plotly)
plot1 <- plot_ly() %>%

 add_trace(data = df %>%
              filter(State == 'Virginia'),x = ~Date , y = ~Janssen.1st.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Janssen Dose') %>%
  
   add_trace(data = df %>%
               
               filter(State == 'Virginia'),x = ~Date , y = ~Pfizer.2nd.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Pfizer 2nd dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Pfizer.1st.Dose.Allocations, type = 'scatter', mode = 'lines+markers',showlegend = TRUE, name = 'Pfizer 1st dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Moderna.2nd.Dose.Allocations, type = 'scatter', mode ='lines+markers',showlegend = TRUE, name = 'Moderna 2nd dose')%>%

  add_trace(data = df %>%
               filter(State == 'Virginia'),x = ~Date , y = ~Moderna.1st.Dose.Allocations, type = 'scatter', mode ='lines+markers',showlegend = TRUE, name = 'Moderna 1st dose')%>%
  layout(
    title = 'Vaccine Allocations from various providers for Viginia',
    xaxis = list(title = 'Date'),
    yaxis = list(title = 'doses allocated')
  )


library(htmlwidgets)
saveWidget(plot1, "Figure1.html", selfcontained = F)
plot1


```



# (b)
```{r}
plot2 <- plot_ly() %>%

add_trace(data = df %>%
filter(State == "Virginia"),
x = ~Date, y = ~Doseperpop, type = 'scatter', mode = 'lines+markers',
showlegend = TRUE, name = 'Virginia dosage allocation over-time' ) %>%
  layout(
    title = 'Vaccine Allocations in Virginia',
    xaxis = list(title = 'Date'),
    yaxis = list(title = 'Number of doses allocated')
  )
  

saveWidget(plot2, "Figure2.html", selfcontained = F)
plot2
```
#Figure 2



#3 District of columbia (D.C) cumulative vacccine allocations;

```{r}
p5 <- plot_ly() %>%
  add_trace(data = df %>%
              filter(State == "District of Columbia"),x = ~Date, y =
              ~Cum.Allocation, type = 'scatter', mode = 'lines+markers',
            showlegend = TRUE, name = 'D.C cumulative allocations'
            ) %>%
  layout(
    title = 'Vaccine Allocations in District of Columbia',
    xaxis = list(title = 'Date'),
    yaxis = list(title = 'Number of doses allocated')
  )
  
p5
```



```{r}

bar_data <- df[, c("State", "Janssen.1st.Dose.Allocations", "Moderna.1st.Dose.Allocations", "Moderna.2nd.Dose.Allocations", "Pfizer.1st.Dose.Allocations", "Pfizer.2nd.Dose.Allocations")]

state_bar_plot <-plot_ly(bar_data, x = ~State, y = ~Janssen.1st.Dose.Allocations, type = 'bar', name = 'Janssen 1st Dose') %>%
  add_trace(y = ~Moderna.1st.Dose.Allocations, name = 'Moderna 1st Dose') %>%
  add_trace(y = ~Moderna.2nd.Dose.Allocations, name = 'Moderna 2nd Dose') %>%
  add_trace(y = ~Pfizer.1st.Dose.Allocations, name = 'Pfizer 1st Dose') %>%
  add_trace(y = ~Pfizer.2nd.Dose.Allocations, name = 'Pfizer 2nd Dose') %>%
  layout(title = 'Total Vaccine Allocations by State',
         xaxis = list(title = 'State'),
         yaxis = list(title = 'Total Allocations'),
         barmode = 'stack')


saveWidget(state_bar_plot, "state_bar_plot.html", selfcontained = F)

state_bar_plot

```


#allocation of vaccine vs population plot
```{r}

avp_plot <- plot_ly(df, x = ~population, y = ~All.Dose.Allocations, text = ~State, mode = 'markers', type = 'scatter', marker = list(size = 10)) %>%
  layout(title = 'Scatter Plot: Population vs. Vaccine Allocations',
         xaxis = list(title = 'Population'),
         yaxis = list(title = 'Vaccine Allocations'))

# Display the plot
avp_plot

saveWidget(avp_plot, "avp_.html", selfcontained = F)

```


```{r}
pie_plot <- plot_ly(df, labels = ~State, values = ~Doseperpop, type = 'pie', hole = 0.4) %>%
  layout(title = 'Proportion of Doses Allocated per Population',
         showlegend = TRUE)

# Display the plot
pie_plot
saveWidget(pie_plot, "pie_plot.html", selfcontained = F)

```


```{r}
library(plotly)
Cum_alloc_plot <- plot_ly(df, x = ~Date, y = ~Cum.Allocation, type = 'scatter', mode = 'lines', line = list(shape = "linear")) %>%
  layout(title = 'Cumulative Vaccine Allocations Over Time',
         xaxis = list(title = 'Date'),
         yaxis = list(title = 'Cumulative Allocations'))

# Display the plot
Cum_alloc_plot
saveWidget(Cum_alloc_plot, "Cum_alloc_plot.html", selfcontained = F)

```



#map(d)
```{r}
install.packages("sp")
install.packages("sf")
install.packages("leaflet")
install.packages("geojsonio")
devtools::install_github("rstudio/leaflet")


```



```{r}
library(geojsonio); library(leaflet); library(dplyr); library(sp); library(sf)

```



```{r}
library(geojsonio)
states <- geojsonio::geojson_read("https://rstudio.github.io/leaflet/json/us-states.geojson", what = "sp")
class(states)
```



## Merging a spatial and non spatial dataframe using sp::merge 

```{r}
states <- sp::merge(states, df, by="name", duplicateGeoms = T)

```




```{r}

m <- leaflet(states) %>%
  setView(-96, 37.8, 4) %>%
  addProviderTiles("MapBox", options = providerTileOptions(
    id = "mapbox.light",
    accessToken = Sys.getenv('MAPBOX_ACCESS_TOKEN')))


```



```{r}
bins <- c(-10, 20, 50, 80, 140, 200)
pal <- colorBin("YlOrRd", domain = states$Doseperpop, bins = bins)

```


```{r}
labels_cases <- sprintf("<strong>%s</strong><br/>Population: %g M<br>
  Dosage per population: %g<br>
  All dose allocation: %g",
  states$name, round(states$population / (1e6), 2),
  states$Doseperpop, states$All.Dose.Allocations) %>%
 lapply(htmltools::HTML)
labels_cases[[1]]
```



```{r}
m <- m %>% addPolygons(
  fillColor = ~pal(states$Doseperpop),
  weight = 2,
  opacity = 1,
  color = "white",
  dashArray = "3",
  fillOpacity = 0.7,
  highlightOptions = highlightOptions(
    weight = 5,
    color = "#666",
    dashArray = "",
    fillOpacity = 0.7,
    bringToFront = TRUE),
  label = labels_cases,
  labelOptions = labelOptions(
    style = list("font-weight" = "normal", padding = "3px 8px"),
    textsize = "15px",
    direction = "auto"))




```


```{r}

m <- m %>% addLegend(pal = pal, values = states$Doseperpop, 
            opacity = 0.7, title = "Dosage per population",
            position = "bottomright")

m
```
#A high Dosageperpop value indicates that particular state performed poorly in allocating the vaccine doses for it's communities.


Figure 4
```{r}
library(htmltools)
library(htmlwidgets)
saveWidget(m, file="m.html", selfcontained = T)
```




